---
title: "MSB1015_Assignment_2"
output: html_document
---
#loading in all the required packages
#Required packages are WikidataQueryServiceR to obtain the data from wikidata
#rJAVA and rcdk are required to predict the pb based on the smiles
#pls to predict bp aswel
```{r setup, include=FALSE}
if(!"RCy3" %in% installed.packages()){
  install.packages("WikidataQueryServiceR")
  install.packages("rJava")
  install.packages ("rcdk") 
  install.packages("pls")
}
  library(WikidataQueryServiceR)
  library(rJava)
  library(rcdk)
  library(pls)
```

#loading in the data using SPARQL
#requesting boiling point of all alkanes with units and SMILES data
```{r}
sparql_query <- 'SELECT DISTINCT ?comp ?compLabel ?bp ?bpUnitLabel ?SMILE WHERE {
    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
  p:P2102 [
    ps:P2102 ?bp ;
    psv:P2102/wikibase:quantityUnit  ?bpUnit
    ] ;
  wdt:P233 ?SMILE .
  SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". }
}'
results = query_wikidata(sparql_query)
```

#data preprocessing and control, remove brackets from smile string for data control
#store SMILES for later use
```{r}
results_SMILES <- results[,5]
pattern<- "\\("
replacement <- ""
results[,5] <- gsub(pattern, replacement, results[,5])
pattern<- "\\)"
results[,5] <- gsub(pattern, replacement, results[,5])
plot(results[order(nchar(results[,5])),3], xlab="alkane length", ylab="bp")
```
#it looks from the plot that the units are not all equal
#changing all of the units kelvin(T(C)+273.15, T(°F) + 459.67) × 5/9)
```{r}
for(i in 1:nrow(results)){
  if (results[i,4] == "degree Celsius"){
  results[i,3] <- results[i,3]+273.15
  results[i,4] <- "kelvin"
}
}
for(i in 1:nrow(results)){
  if (results[i,4] == "degree Fahrenheit"){
  results[i,3] <- ((results[i,3]+ 459.67)*5/9)
  results[i,4] <- "kelvin"
}
}
```
#data  control after changing temperatures to celsius
```{r}
plot(results[order(nchar(results[,5])),3], xlab="alkane length", ylab="bp")

```
#pls for bp prediction on alkanes
